Consistent Polyhedral Surrogates via Embeddings

### 2019-05-16 neurips-19/poly-embed.tex
pass over examples, add lovasz

\begin{definition}[Consistency]\label{def:consistency}
We say that a surrogate $L(u)$ is \emph{consistent} with respect to a discrete loss $\ell(r)$ if, for all $p \in \simplex$, \jessie{....?}
\end{definition}

@@

We omit the details connecting this definition to consistency, and instead refer the reader to Ramaswamy et al.~\cite{ramaswamy2018consistent}.
(See the proof of Theorems 8.1-8.3.).
\raf{We could flesh this out if we want.  Might be worth it.  Their entire proof(s) establish $\delta$-separation, and then the line ``The Theorem simply follows from linearity of expectation.''}

Several consistent convex surrogates in the literature can be thought of as ``embeddings'': map the discrete reports to a vector space, and find a convex loss which agrees with the original loss and has the same statistical behavior.
We formalize this notion as follows.
%
\begin{definition}\label{def:loss-embed}
  A loss $L:\reals^d\to\reals^\Y$ \emph{embeds} a loss $\ell:\R\to\reals^\Y$ if there exists some injective embedding $\varphi:\R\to\reals^d$ such that
  (i) for all $r\in\R$ we have $L(\varphi(r)) = \ell(r)$, and (ii) for all $p\in\simplex,r\in\R$ we have
  \begin{equation}\label{eq:embed-loss}
    r \in \argmin_{r'\in\R} \inprod{p}{\ell(r')} \iff \varphi(r) \in \argmin_{u\in\reals^d} \inprod{p}{L(u)}~.
  \end{equation}
\end{definition}
%
Note that embeddings do not automatically give rise to separated links; indeed, apart from mapping the embedded reports back from whence they came \raft{I know... just wanted to say that} via $\psi(\varphi(r)) = r$, how to map the remaining values is far from clear.
Constructing separated links from embeddings will therefore be a major focus in Section~\ref{sec:poly-loss-embed}.

\raf{How hinge and abstain fit in}
\jessie{Added the paragraphs below.}
To illustrate the idea of embeddings, we now look in detail at two key examples.
First, consider a binary classification problem on the outcomes $\{-1, +1\}$.
We know that both the hinge loss $L_{hinge}(u,y) := (1 - uy)_+$ and logistic loss $L_{logistic}(u,y) := \ln\left(1 + \exp(-uy)\right)$ are convex surrogates that are minimized in expectation by predicting the mode, as is 0-1 loss $\ell_{mode}(r,y) := \textbf{1}[r\neq y]$.
However, we say that hinge loss embeds the mode, while logistic loss does not.
\jessie{Come back to flesh this out if space?}

In another example, we consider the abstain loss and an embedding given by Ramaswamy et al.~\cite{ramaswamy2018consistent} in Section~\ref{sec:examples}, but note that the link function they provide is not unique.
Essentially, the discrete abstain$(1/2)$ loss is minimized in expectation by predicting the most likely outcome $\argmax_y p_y$, conditioned on $p_y \geq 1/2$.
If the most likely outcome is not likely enough, the loss is minimized by ``abstaining'' from predicting; a situation one can imagine would be useful if a false positive is costly.
\jessie{Are we adding discussion about the choice of link relative to the norm?}

@@

\subsection{Important Examples}\label{sec:important-examples}
\raf{Probably should move this up to where we introduce embeddings.}

  Ranking and classification problems are well-known to the field of machine learning.
  The most studied classification problem is the elicitation of the mode.
  While the mode over $n$ outcomes is elicitable by 0-1 loss, it is known that eliciting the mode by a convex loss is much more difficult.
  In fact, the best convex surrogate that can be used to elicit the mode ``embeds'' the reports into $\reals^{n-1}$.
  Therefore, we say $\eliccvx($mode$) = n-1$.

  A property of interest similar to the mode is the $\abstain{\alpha}$ property.
  In this property in which one reports the most likely outcome if the probability of such an outcome is greater than $1-\alpha$, and otherwise ``abstains'' from reporting.
  That is, $\gamma(p) = \argmax_{r \in \R}p_r$ if $\max_{r \in \R} p_r \geq 1-\alpha$ and is ``abstain'' otherwise.
  The finite loss eliciting $\abstain{\alpha}$ can be written
  $\ell(r,y) = 0$ if $r=y$, $\alpha$ if $r=\text{abstain}$, and $1$ otherwise.
\cite{ramaswamy2018consistent} show that $\elicpoly(\abstain{1/2}) \leq \ceil{\log_2 n}$, although it is not known if this bound is tight.

@@

@@

\begin{proposition}
  If $\psi$ is a separated link between losses $L$ and $\ell$, then \raf{how do we want to talk about consistency?}
\end{proposition}
\begin{proof}
  \raf{Use Hari's theorems from their section 8.}
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to specific surrogates}\label{sec:examples}

Our results give a framework to construct consistent surrogates and link functions, but they also provide a way to verify the consistency or inconsistency of given surrogates.
Below, we illustrate the power of this framework with specific examples from the literature, as well as new examples.
In some cases we simplify existing proofs, while in others our results we give new results, such as the inconsistency of the recently proposed Lov\'asz hinge.
\raf{Notes to self: The top-k and hypercube/set examples (except abstain) have the link built in, e.g.\ as the sign function), and they just work around it.  Our results may suggest that it's worth thinking ``outside the box'' (HAH, genious...) and looking for embeddings which are not the hypercube.}

@@

\subsection{Inconsistency of top-$k$ loss}
Yang and Koyejo~\cite{yang2018consistency} prove the inconsistency of a family of surrogates for the top-$k$ classification problem.

The top $k$ classification problem takes as input a vector $r \in \R = \{r \in \{0,1\}^n \text{ such that } |r| = k\}$, and observes outcome $y \in [n]$

@@

\subsection{Inconsistency of Lov\'asz hinge}
\raf{NEW: no consistency results known previously}

\raf{Will not even define the hinge, just mention informal, and give the expression explicitly for the increasing k=2 case: max f(1)... whatever. (That way it can fit in a few paragraphs, and justifies the delegation of the full result to the full version.)}

\jessie{Lovasz hinge paper~\cite{yu2018lovasz}}
\raf{I'm tempted to do the following.  We give the abstain counter-example for $k=2$ (I've been using $k$ for the number of elements in the set constructions, but maybe $n$ is better; we can sync up notation later), and then characterize the $k=2$ case informally, showing that the only thing that works is when $f(\{1,2\}) = f(\{1\}) + f(\{2\})$.  Then we can say ``In fact, in full version of the paper, we show a much tighter result: essentially the only Lov\'asz hinge that is consistent is weighted Hamming loss'' or something to that effect.  This leaves room for us to send a journal version to JMLR (for which we'll need some new results), which seems like a good idea for a paper like this: We seem to have useful tools here, so why not let them marinate a bit in NeurIPS, collect feedback, and then publish a more polished/mature version.}

>>>>

\begin{definition}[Consistency]\label{def:consistency}
  A surrogate $L:\reals^d\to\reals^\Y_+$ and link $\psi:\reals^d\to\R$ is \emph{consistent} with a discrete loss $\ell:\R\to\reals^\Y_+$ if \raf{statement about distributions on $\X\times\Y$, limits of hypotheses, and Bayes opt hypotheses}
\end{definition}

@@

Several consistent convex surrogates in the literature can be thought of as ``embeddings'', wherein one maps the discrete reports to a vector space, and finds a convex loss which agrees with the original loss and has the same statistical behavior.
We formalize this notion as follows.
%
\begin{definition}\label{def:loss-embed}
  A loss $L:\reals^d\to\reals^\Y$ \emph{embeds} a loss $\ell:\R\to\reals^\Y$ if there exists some injective embedding $\varphi:\R\to\reals^d$ such that
  (i) for all $r\in\R$ we have $L(\varphi(r)) = \ell(r)$, and (ii) for all $p\in\simplex,r\in\R$ we have
  \begin{equation}\label{eq:embed-loss}
    r \in \argmin_{r'\in\R} \inprod{p}{\ell(r')} \iff \varphi(r) \in \argmin_{u\in\reals^d} \inprod{p}{L(u)}~.
  \end{equation}
\end{definition}
%
Note that it is not clear if embeddings give rise to separated links; indeed, apart from mapping the embedded reports back from whence they came \raft{I know... just wanted to say that} via $\psi(\varphi(r)) = r$, how to map the remaining values is far from clear.
Constructing separated links from embeddings will therefore be a major focus in Section~\ref{sec:poly-loss-embed}.

\raf{How hinge fits in}
\jessie{Added the paragraphs below.}
To illustrate the idea of embeddings, we now look in detail at two key examples.
First, consider a binary classification problem on the outcomes $\{-1, +1\}$.
We know that both the hinge loss $L_{hinge}(u,y) := (1 - uy)_+$ and logistic loss $L_{logistic}(u,y) := \ln\left(1 + \exp(-uy)\right)$ are convex surrogates that are minimized in expectation by predicting the mode, as is 0-1 loss $\ell_{mode}(r,y) := \ones\{r\neq y\}$.
However, we say that hinge loss embeds the mode, while logistic loss does not.
\jessie{Come back to flesh this out if space?}

% In another example, we consider the abstain loss and an embedding given by Ramaswamy et al.~\cite{ramaswamy2018consistent} in Section~\ref{sec:examples}, but note that the link function they provide is not unique.
% Essentially, the discrete abstain$(1/2)$ loss is minimized in expectation by predicting the most likely outcome $\argmax_y p_y$, conditioned on $p_y \geq 1/2$.
% If the most likely outcome is not likely enough, the loss is minimized by ``abstaining'' from predicting; a situation one can imagine would be useful if a false positive is costly.

@@

@@

We have now seen the tight relationship between polyhedral losses and embeddings; in particular, every polyhedral loss embeds some discrete loss.
The embedding itself tells us how to map the embedded points back to the discrete reports (map $\varphi(r)$ to $r$), but it is not clear how to map the remaining reports to obtain a full link function, and whether such a link can lead to consistency.
In this section, we show that separated link functions guarantee consistency, and give a construction to generate separated links for any polyhedral loss.
We illustrate our construction in the following section.

@@

\begin{proposition}
  If $\psi$ is a separated link between losses $L$ and $\ell$, then $L$ and $\psi$ are consistenty with $\ell$.
\end{proposition}
\begin{proof}
  \raf{Use Hari's theorems from their section 8.}
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to Specific Surrogates}\label{sec:examples}

Our results give a framework to construct consistent surrogates and link functions, but they also provide a way to verify the consistency or inconsistency of given surrogates.
Below, we illustrate the power of this framework with specific examples from the literature, as well as new examples.
In some cases we simplify existing proofs, while in others we give new results, such as the inconsistency of the recently proposed Lov\'asz hinge.
\raf{Notes to self: The top-k and hypercube/set examples (except abstain) have the link built in, e.g.\ as the sign function), and they just work around it.  Our results may suggest that it's worth thinking ``outside the box'' (HAH, genious...) and looking for embeddings which are not the hypercube.}

@@

\subsection{Inconsistency of top-$k$ losses}

Yang and Koyejo~\cite{yang2018consistency} prove the inconsistency of a family of polyhedral surrogates for the top-$k$ classification problem.
To illustrate our framework, we show how to simplify their arguments using embeddings.

The top $k$ classification problem takes as input a vector $r \in \R = \{r \in \{0,1\}^n \text{ such that } |r| = k\}$, \raf{mixing sets and indicators here, unless you mean $\|r\|_1$} and observes outcome $y \in [n]$

@@

\subsection{Inconsistency of Lov\'asz hinge}
\label{sec:lovasz-hinge}

Many structured prediction settings can be thought of as making multiple predictions at once, with a loss function that jointly measures error based on the relationship between these predictions.\raf{CITE; can't remember which}
In the case of $k$ binary predictions, these settings are typically formalized by taking the predictions and outcomes to be $\pm 1$ vectors, so $\R=\Y=\{-1,1\}^k$.
One then defines a joint loss function, which is often merely a function of the set of mispredictions, meaning $\ell^g(r)_y = g(\{i \in [k] : r_i \neq y_i\})$ for some function $g:2^{[k]}\to\reals$.
In an effort to provide a general convex surrogate for these settings when $g$ is a submodular function, Yu and Blaschko~\cite{yu2018lovasz} introduce the \emph{Lov\'asz hinge}, which leverages the well-known convex Lov\'asz extension of submodular functions.
While the authors provide theoretical justification and experiments, consistency of the Lov\'asz hinge is left open, which we resolve negatively.

Rather than formally define the Lov\'asz hinge, we defer the full analysis to the full version of the paper\raf{CITE -- arXiv}, and focus here on the $k=2$ case.
For brevity, we write $g_\emptyset := g(\emptyset)$, $g_{1,2} := g(\{1,2\})$, etc.
Assuming $g$ is normalized and increasing (meaning $g_{1,2} \geq \{g_1,g_2\} \geq g_\emptyset = 0$), the Lov\'asz hinge $L:\reals^k\to\reals^\Y_+$ is given by
\begin{multline}
  \label{eq:lovasz-hinge}  
  L^g(u)_y = \max\Bigl\{(1-u_1y_1)_+ g_1 + (1-u_2y_2)_+ (g_{1,2}-g_1),\\[-4pt] (1-u_2y_2)_+ g_2 + (1-u_1y_1)_+ (g_{1,2}-g_2)\Bigr\}~,
\end{multline}
where $(x)_+ = \max\{x,0\}$.
We will explore the range of values of $g$ for which $L^g$ is consistent, where the link function $\psi:\reals^2\to\{-1,1\}^2$ is fixed as $\psi(u)_i = \sgn(u_i)$, with ties broken arbitrarily.

Let us consider $g_\emptyset = 0$, $g_1 = g_2 = g_{1,2} = 1$, for which $\ell^g$ is merely 0-1 loss on $\Y$.
For consistency, then, for any distribution $p\in\simplex$, we must have that whenever $u \in \argmin_{u'\in\reals^2} p\cdot L^g(u')$, then $\psi(u)$ must be the most likely outcome, in $\argmax_{y\in\Y} p(y)$.
Simplifying eq.~\eqref{eq:lovasz-hinge}, however, we have
\begin{equation}
  \label{eq:lovasz-hinge-abstain}  
  L^g(u)_y = \max\bigl\{(1-u_1y_1)_+,(1-u_2y_2)_+\bigr\} = \max\bigl\{1-u_1y_1,1-u_2y_2,0\bigr\}~,
\end{equation}
which is exactly the abstain surrogate\raf{eqref here}.
We immediately conclude that $L^g$ cannot be consistent with $\ell^g$, as the origin will be the unique optimal report for $L^g$ under distributions with $p_y < 0.5$ for all $y$, and one can simply take a distribution which disagrees with the way ties are broken in $\psi$.
For example, if we take $\sgn(0) = 1$, then under $p((1,1)) = p((1,-1)) = p((-1,1)) = 0.2$ and $p((-1,-1)) = 0.4$, we have $\{0\} = \argmin_{u\in\reals^2} p\cdot L^g(u)$, yet $\psi(0) = (1,1) \notin \{(-1,-1)\} = \argmin_{r\in\R} p\cdot\ell^g(r)$.

In fact, this example is typical: one can show that $L^g$ is consistent if and only if $g_{1,2} = g_1 + g_2$.
Moreover, in the linear case, which corresponds to $g$ being \emph{modular}, the Lov\'asz hinge reduces to weighted Hamming loss, which is known to be consistent.
In the full version of the paper\raf{CITE again arXiv} we generalize this observation for all $k$: $L^g$ is consistent if and only if $g$ is modular.
In other words, even for $k>2$, the only consistent Lov\'asz hinge is weighted Hamming loss.
These results cast doubt on the effectiveness of the Lov\'asz hinge in practice.
